---
title: "Interpreting results with azul"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interpreting results with azul}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(azul)
```

`azul` converts a fitted model, a hypothesis test, a result table, or a raw
dataset into a manuscript-ready interpretation in MJMS / DrPH style. The single
entry point is `interpret()`.

## Descriptive statistics

```{r}
interpret(mtcars)                       # Table 1 style narrative
```

## A hypothesis test

```{r}
interpret(t.test(mpg ~ am, data = mtcars))
```

## Linear regression

Continuous predictors and factor levels are worded differently, and the model
fit (R-squared, overall F) is reported.

```{r}
d <- transform(mtcars, gear = factor(gear))
interpret(lm(mpg ~ wt + gear, d), outcome = "fuel economy", unit = "mpg")
```

## Logistic regression

```{r}
d <- transform(mtcars, am = factor(am))
interpret(glm(am ~ wt + hp, d, family = binomial), outcome = "manual transmission")
```

## Survival (Cox)

```{r}
library(survival)
interpret(coxph(Surv(time, status) ~ age + sex, lung), outcome = "death")
```

## Interaction stratification

When an interaction is statistically significant, `azul` reports the stratified
simple effects instead of the marginal main effects.

```{r}
set.seed(1); n <- 300
d <- data.frame(sex = factor(sample(c("F","M"), n, TRUE)),
                trt = factor(sample(c("placebo","drug"), n, TRUE)))
d$y <- 10 + 3*(d$trt=="drug") + 8*(d$trt=="drug")*(d$sex=="M") + rnorm(n, 0, 4)
interpret(lm(y ~ sex * trt, d), outcome = "symptom score")
```

## A results table

```{r}
tab <- data.frame(term = c("Male","Age"), OR = c(1.49, 1.03),
                  LL = c(1.24, 1.01), UL = c(1.78, 1.05), p = c(4e-4, 0.02))
interpret_table(tab, type = "OR", outcome = "coronary artery disease")
```

## Diagnostic accuracy

```{r}
tb <- matrix(c(85, 15, 20, 180), 2, byrow = TRUE)
interpret_diagnostic(tb, test = "the rapid test")
```

## Manuscript paragraph vs console layout

`print()` (the default) shows a spaced, sub-headed layout. Use `as.character()`
to get the single flowing paragraph for pasting into a thesis.

```{r}
x <- interpret(lm(mpg ~ wt, mtcars), outcome = "fuel economy")
cat(as.character(x))
```
